Local PCA results

### 2016-05-11 medicago/medicago_data_setup.Rmd
run_on_medicago.R runs, saves regions and labeled MDS coords

```{r setup}
devtools::load_all("../package")
library(Matrix)
library(colorspace)
options(width=100)
```

@@

To get windows out of the VCF file and recode them, we'll just pipe a call to `bcftools`.
Here's a dummy version that will pull out windows of 100 SNPs for only the first four individuals.
```{r win_fn}
get.indivs <- c("HM001","HM002","HM003","HM004")
bcf.file <- "data/chr1-filtered-set-2014Apr15.bcf"
# first read in the info about sites
bcf.con <- pipe(paste("bcftools query -f '%POS\\n' -r chr1:1-100000",bcf.file),open="r")
bcf.sites <- list( chr1=scan(bcf.con) )
close(bcf.con)
# now use bcftools to pull out sites we want
win.fn <- function (n,win.snps=100) {
    sites <- list( chr1=bcf.sites[[1]][seq(1+(n-1)*win.snps,length.out=win.snps)] )
    regions <- enclosing_region(sites)
    query_genotypes( bcf.file, regions, get.indivs )
}
win.fn(20,10)  # get the 20th window of 10 SNPs
```

Here's an alternate version that pulls out windows based on basepairs.
Note that this version assumes we have only one chromosome, starting at zero.
```{r win_fn_bp}
win.fn.bp <- function (n,win.bp=200) {
    win.start <- 1+(n-1)*win.bp
    win.end <- win.start + win.bp-1
    regions <- data.frame( chrom="chr1", start=win.start, end=win.end )
    query_genotypes( bcf.file, regions, get.indivs )
}
win.fn.bp(20,50)
```

@@

Now let's extract the extreme windows in the MDS plot:
```{r get_corners}
mincirc <- enclosing_circle( mds.coords[,1:2] )
mds.corners <- corners( mds.coords[,1:2], prop=.05 )
corner.cols <- c("red","blue","purple")
ccols <- rep("black",nrow(mds.coords))
for (k in 1:ncol(mds.corners)) {
    ccols[ mds.corners[,k] ] <- corner.cols[k]
}
plot( mds.coords[,1:2], pch=20, col=adjustcolor(ccols,0.75), 
        xlab="MDS coordinate 1", ylab="MDS coordinate 2", 
        xlim=mincirc$ctr[1]+c(-1,1)*mincirc$rad,
        ylim=mincirc$ctr[2]+c(-1,1)*mincirc$rad )
plot_circle( mincirc, col='red' )
points( mincirc$three, col='red', cex=2 )
points( mds.coords[mincirc$index,], col='red', cex=1.5 )
```

>>>>

```{r setup}
devtools::load_all("../package")
library(Matrix)
library(colorspace)
options(width=100)
fig.dim <- 4
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.height=fig.dim,fig.align='center')
```

```

@@

To get windows out of the VCF file and recode them, we'll just pipe a call to `bcftools`.
Here's a dummy version that will pull out windows of 100 SNPs for only the first four individuals.
```{r win_fn}
get.indivs <- c("HM001","HM002","HM003","HM004")
bcf.file <- "data/chr1-filtered-set-2014Apr15.bcf"
# first read in the info about sites
bcf.con <- pipe(paste("bcftools query -f '%POS\\n' -r chr1:1-100000",bcf.file),open="r")
bcf.sites <- list( chr1=scan(bcf.con) )
close(bcf.con)
# now use bcftools to pull out sites we want
win.fn <- function (n,win.snps=100) {
    sites <- list( chr1=bcf.sites[[1]][seq(1+(n-1)*win.snps,length.out=win.snps)] )
    regions <- enclosing_region(sites)
    vcf_query( bcf.file, regions, get.indivs )
}
win.fn(20,10)  # get the 20th window of 10 SNPs
```

Here's an alternate version that pulls out windows based on basepairs.
Note that this version assumes we have only one chromosome, starting at zero.
```{r win_fn_bp}
win.fn.bp <- function (n,win.bp=200) {
    win.start <- 1+(n-1)*win.bp
    win.end <- win.start + win.bp-1
    regions <- data.frame( chrom="chr1", start=win.start, end=win.end )
    vcf_query( bcf.file, regions, get.indivs )
}
win.fn.bp(20,50)
```

@@

Now let's extract the extreme windows in the MDS plot:
```{r get_corners}
mincirc <- lostruct:::enclosing_circle( mds.coords[,1:2] )
mds.corners <- corners( mds.coords[,1:2], prop=.05 )
corner.cols <- c("red","blue","purple")
ccols <- rep("black",nrow(mds.coords))
for (k in 1:ncol(mds.corners)) {
    ccols[ mds.corners[,k] ] <- corner.cols[k]
}
plot( mds.coords[,1:2], pch=20, col=adjustcolor(ccols,0.75), 
        xlab="MDS coordinate 1", ylab="MDS coordinate 2", 
        xlim=mincirc$ctr[1]+c(-1,1)*mincirc$rad,
        ylim=mincirc$ctr[2]+c(-1,1)*mincirc$rad )
plot_circle( mincirc, col='red' )
points( mincirc$three, col='red', cex=2 )
points( mds.coords[mincirc$index,], col='red', cex=1.5 )
```

### 2016-05-11 medicago/summarize_run.Rmd
run_on_medicago.R runs, saves regions and labeled MDS coords

```{r setup, include=FALSE}
library(lostruct)
library(colorspace)
library(jsonlite)
opt <- fromJSON("config.json")

    pca.file <- file.path( opt$outdir, sprintf(gsub(".bcf",".pca.csv",basename(bcf.file))) )
    if (file.exists(pca.file)) { stop(paste("File",pca.file,"already exists! Not overwriting.")) }
    cat("Finding PCs for", bcf.file, "and writing out to", pca.file, "\n")
    win.fn <- vcf_windower(bcf.file, size=opt$size, type=tolower(opt$type) )

win.regions <- region(win.fn.snp)()
win.mids <- (win.regions$start+win.regions$end)/2
chrom.cols <- rainbow_hcl(128, c=90, end=.9*360)[as.numeric(cut(win.mids,128))]
pairs( mds.coords, pch=20, col=adjustcolor(chrom.cols,0.75) )

```

>>>>

```{r setup, include=FALSE}
library(lostruct)
library(colorspace)
library(jsonlite)
fig.dim <- 4
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.height=fig.dim,fig.align='center')
```

```{r data_setup, include=FALSE}
opt <- fromJSON("config.json")

# data files, precomputed
pca.files <- list.files(".",".*.pca.csv")
mds.file <- "mds_coords.csv"
regions.file <- gsub("[.]pca[.]csv$",".regions.csv",pca.files)

# read in mds
mds.coords <- read.csv(mds.file,header=TRUE)
mds.cols <- (1:ncol(mds.coords))[-(1:2)]

# position information
regions <- do.call( rbind, lapply( regions.file, read.csv, header=TRUE ) )
regions$mid <- (regions$start+regions$end)/2
chrom.cols <- rainbow_hcl(128, c=90, end=.9*360)[as.numeric(cut(regions$mid,128))]
```

This run had these parameters:

- *window type* : `r opt$type`
- *window size* : `r opt$size`
- *number of pcs* : `r opt$npc`
- *number of MDS coordinates* : `r opt$nmds`

Here are the computed MDS coordinates, colored by chromosome:
```{r mds_pairplot}
pairs( mds.coords[,mds.cols], pch=20, col=adjustcolor(chrom.cols,0.75) )
```

Here are the extreme windows in the MDS plot:
```{r get_corners}
mds.corners <- corners( mds.coords[,mds.cols[1:2]], prop=.05 )
corner.cols <- c("red","blue","purple")
ccols <- rep("black",nrow(mds.coords))
for (k in 1:ncol(mds.corners)) {
    ccols[ mds.corners[,k] ] <- corner.cols[k]
}

plot( mds.coords[,mds.cols[1:2]], pch=20, col=adjustcolor(ccols,0.75), 
        xlab="MDS coordinate 1", ylab="MDS coordinate 2" )

for (k in mds.cols) {
    plot( regions$mid/1e6, mds.coords[,k], pch=20, 
            xlab="Position (Mb)", ylab=paste("MDS coordinate",k),
            col=adjustcolor(ccols,0.75) )
}
```
